How Short Lived is the Population III Epoch?

### 2010-09-07 ms.tex
Adding stellar feedback section and completing star formation section.

We use the adaptive mesh refinement (AMR) code
\enzo~\citep{BryanNorman97, OShea2004} that employs our implementation
of adaptive ray tracing for radiation transport \citep{Wise10}.  To
resolve minihalos with at least 100 dark matter (DM) particles and
follow the formation of the first generation of dwarf galaxies, we use
a simulation box of 1 Mpc that has a resolution of $256^3$.  This
gives us a DM mass resolution of 1840 \Ms.  We refine the grid on
baryon overdensities of $3 \times 2^{-0.2l}$, where $l$ is the AMR
level, resulting in a super-Lagrangian behavior.  We also refine on a
DM overdensity of $3 \times 2^l$ and always resolve the local Jeans
length by at least 4 cells.  The latter criterion avoids any artifical
fragmentation during gaseous collapses \citep{Truelove97}.  We
initialize the simulation with \textsl{grafic} \citep{Bertschinger01}
at $z = ???$ and use the cosmological parameters from the 7-year WMAP
data \citep{ref}: $\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$,
$\Omega_b = 0.0449$, $h = 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$
with the variables having their usual definitions.

@@

For Pop III stars, we use the same star formation model as
\citet{Abel07} and \citet{Wise08_Gal} where each star particle
represents a single star.  Instead of using a fixed stellar mass, we
randomly sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we initialize both
simulations with the same random seed and record the number of times
the random number generator \citep[Mersenne twister;][]{ref} has been
called for use when restarting the simulations.  Although the Pop III
star formation history (SFH) is identical for a given $N$ stars, the
Pop III SFH for a given halo can be different between the two
simulations.  This happens because the order in which halos form Pop
III stars can be altered from enhanced star formation and feedback
from nearby Pop II star clusters, as we will demonstrate in this
Letter.

@@

>>>>

We use the adaptive mesh refinement (AMR) code
\enzo~\citep{BryanNorman97, OShea2004} that employs our implementation
of adaptive ray tracing for radiation transport \citep{Wise10}.  To
resolve minihalos with at least 100 dark matter (DM) particles and
follow the formation of the first generation of dwarf galaxies, we use
a simulation box of 1 Mpc that has a resolution of $256^3$.  This
gives us a DM mass resolution of 1840 \Ms.  We refine the grid on
baryon overdensities of $3 \times 2^{-0.2l}$, where $l$ is the AMR
level, resulting in a super-Lagrangian behavior.  We also refine on a
DM overdensity of $3 \times 2^l$ and always resolve the local Jeans
length by at least 4 cells.  The latter criterion avoids any artifical
fragmentation during gaseous collapses \citep{Truelove97}.  We
initialize the simulation with \textsl{grafic} \citep{Bertschinger01}
at $z = 130$ and use the cosmological parameters from the 7-year WMAP
data \citep{WMAP7}: $\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$,
$\Omega_b = 0.0449$, $h = 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$
with the variables having their usual definitions.

@@

For Pop III stars, we use the same star formation model as
\citet{Abel07} and \citet{Wise08_Gal} where each star particle
represents a single star.  The critical overdensity in which a star
forms is $5 \times 10^5$.  Instead of using a fixed stellar mass, we
randomly sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we initialize both
simulations with the same random seed and record the number of times
the random number generator \citep[Mersenne twister;][]{ref} has been
called for use when restarting the simulations.  Although the Pop III
star formation history (SFH) is identical for a given $N$ stars, the
Pop III SFH for a given halo can be different between the two
simulations.  This happens because the order in which halos form Pop
III stars can be altered from enhanced star formation and feedback
from nearby Pop II star clusters, as we will demonstrate in this
Letter.

We treat Pop II star formation with the same prescription as
\citet{Wise09}, which is a modified version of the widely used
\citet{Cen92} method but accounts for the fact that the molecular
clouds are resolved.  The critical overdensity is the same as the Pop
III star formation model.  In each star-forming region, a fraction
$c_* = 0.07$ of the cold gas ($T < 10^3$ K) is removed from the grid
and deposited into the star particle.  Each star particle lives for 20
Myr, the lifetime of an OB star.  These stars generate the majority of
the ionizing radiation and SNe feedback in stellar clusters, thus we
ignore lower mass stars.  We discuss our treatment of stellar feedback
next.

@@

The mass-dependent luminosities and lifetimes of the Pop III stars are
taken from \citet{Schaerer02}.  The radiation field is evolved with
with adaptive ray tracing \citep{Abel02, Wise10} and is coupled
self-consistently to the hydrodynamics.  These stars die as
pair-instability SNe (PISNe) if they are in the mass range 140--260
\Ms \citep{Heger02}.  We use the explosion energy from
\citeauthor{Heger02}, where we fit the following function to their
models, $E_{\rm PISN} = 10^{51} \times [5.0 + 1.304 (M_{\rm He} -
64)]$, where $M_{\rm He} = (13/24) \times (M_\star - 20) \Ms$ is the
helium core mass and $M_\star$ is the stellar mass.

The Pop II stars emit 6000 hydrogen ionizing photons per baryon over
their lifetime, and we do not consider singly- and doubly-ionizing
helium photons.  We note that low-metallicity stars generate up to a
factor of four more ionizing photons than a solar metallicity
population \citep{Schaerer03} and might be underestimating the
radiative feedback.  Nonetheless this study provides an excellent
first insight in the transition to Pop II star formation, as the metal
enrichment is the key ingredient here.  For SN feedback, these stars
generate $6.8 \times 10^{48}$ erg s$^{-1}$ $\Ms^{-1}$ of energy after
living for 4 Myr, which is injected into spheres with a radius of 10
pc.  If the resolution of the grid is less than 10/3 pc, we deposit
the energy into a $3^3$ cube surrounding the star particle.
